---
title: "B Call Modeling Template"
output: html_document
---

## Load Packages and Data

```{r, message = F, warning = F}
# Load in the libraries
library(tidyverse)
library(sdmTMB)
library(ggplot2)
library(janitor)
library(sp)
library(readr)
library(ggcorrplot)
library(Metrics)
library(naniar)
library(tibble)
library(rsample)
library(caret)
library(purrr)
library(tune)

# Read in the data
cal_data <- read.csv("../../data/Merged_CalCOFI/STD_CalCOFI_final")
```

## Data Cleaning

```{r}
# Removing the first two columns
cal_data <- cal_data[, -c(1, 2)]

# Removing larvae and euphasia abundance
cal_data <- cal_data %>% 
  select(-euphasia_abundance, -euphasia_abundance_std, -larvae_count, -larvae_count_std, - total_abundance, -total_abundance_std)

# Changing dataset to only have present response data
cal_data <- cal_data %>% 
  filter(!is.na(b_scaled))
```

## Exploratory Data Analysis

### Relationship Between Response and Predictors

#### B Calls vs. Longitude

```{r, message = FALSE}
# Scatterplot with line of fit
ggplot(cal_data, aes(x = lon, y = b_scaled)) +
  geom_point() +                 # Scatter points
  geom_smooth(method = "lm")  +    # Linear model fit line
  theme_classic()
```

#### B Calls vs. Latitude

```{r, message = FALSE}
# Scatterplot with line of fit
ggplot(cal_data, aes(x = lat, y = b_scaled)) +
  geom_point() +                 # Scatter points
  geom_smooth(method = "lm")  +    # Linear model fit line
  theme_classic()
```

#### B Calls vs. Date

```{r, message = FALSE}
# Scatterplot with line of fit
ggplot(cal_data, aes(x = date, y = b_scaled)) +
  geom_point() +                 # Scatter points
  theme_classic()
```

#### B Calls vs. Season

```{r, message = FALSE}
# Scatterplot with line of fit
ggplot(cal_data, aes(x = season, y = b_scaled)) +
  geom_point() +                 # Scatter points
  geom_smooth()  +    # Linear model fit line
  theme_classic()
```

#### B Calls vs. Estimated Depth

```{r, message = FALSE}
# Scatterplot with line of fit
ggplot(cal_data, aes(x = est_depth, y = b_scaled)) +
  geom_point() +                 # Scatter points
  geom_smooth(method = "lm")  +    # Linear model fit line
  theme_classic()
```

#### B Calls vs. Larvae Count

```{r, warning = FALSE, message = FALSE}
# Scatterplot with line of fit
ggplot(cal_data, aes(x = larvae_count, y = b_scaled)) +
  geom_point() +                 # Scatter points
  geom_smooth(method = "lm")  +    # Linear model fit line
  theme_classic()
```

#### B Calls vs. Temperature at Depth 55

```{r, message = FALSE}
# Scatterplot with line of fit
ggplot(cal_data, aes(x = temperature_depth_55, y = b_scaled)) +
  geom_point() +                 # Scatter points
  geom_smooth(method = "lm") + # Linear model fit line
  theme_classic()
```

#### B Calls vs. Temperature at Depth 105

```{r, message = FALSE}
# Scatterplot with line of fit
ggplot(cal_data, aes(x = temperature_depth_105, y = b_scaled)) +
  geom_point() +                 # Scatter points
  geom_smooth(method = "lm") + # Linear model fit line
  theme_classic()
```

#### B Calls vs. Temperature at Depth 280

```{r, message = FALSE}
# Scatterplot with line of fit
ggplot(cal_data, aes(x = temperature_depth_280, y = b_scaled)) +
  geom_point() +                 # Scatter points
  geom_smooth(method = "lm") + # Linear model fit line
  theme_classic()
```

#### B Calls vs. Pressure at Depth 55

```{r, message = FALSE}
# Scatterplot with line of fit
ggplot(cal_data, aes(x = pressure_depth_55, y = b_scaled)) +
  geom_point() +                 # Scatter points
  geom_smooth(method = "lm") + # Linear model fit line
  theme_classic()
```

#### B Calls vs. Pressure at Depth 105

```{r, message = FALSE}
# Scatterplot with line of fit
ggplot(cal_data, aes(x = pressure_depth_105, y = b_scaled)) +
  geom_point() +                 # Scatter points
  geom_smooth(method = "lm") + # Linear model fit line
  theme_classic()
```

#### B Calls vs. Pressure at Depth 280

```{r, message = FALSE}
# Scatterplot with line of fit
ggplot(cal_data, aes(x = pressure_depth_280, y = b_scaled)) +
  geom_point() +                 # Scatter points
  geom_smooth(method = "lm") + # Linear model fit line
  theme_classic()
```

#### A Calls vs. Salinity at Depth 55

```{r, message = FALSE}
# Scatterplot with line of fit
ggplot(cal_data, aes(x = salinity_depth_55, y = b_scaled)) +
  geom_point() +                 # Scatter points
  geom_smooth(method = "lm") + # Linear model fit line
  theme_classic()
```

#### B Calls vs. Salinity at Depth 105

```{r, message = FALSE}
# Scatterplot with line of fit
ggplot(cal_data, aes(x = salinity_depth_105, y = b_scaled)) +
  geom_point() +                 # Scatter points
  geom_smooth(method = "lm") + # Linear model fit line
  theme_classic()
```

#### B Calls vs. Salinity at Depth 280

```{r, message = FALSE}
# Scatterplot with line of fit
ggplot(cal_data, aes(x = salinity_depth_280, y = b_scaled)) +
  geom_point() +                 # Scatter points
  geom_smooth(method = "lm") + # Linear model fit line
  theme_classic()
```

#### B Calls vs. Magnitude at Depth 55

```{r, message = FALSE, warning = FALSE}
# Scatterplot with line of fit
ggplot(cal_data, aes(x = magnitude_depth_55, y = b_scaled)) +
  geom_point() +                 # Scatter points
  geom_smooth(method = "lm") + # Linear model fit line
  theme_classic()
```

#### B Calls vs. Magnitude at Depth 105

```{r, message = FALSE, warning = FALSE}
# Scatterplot with line of fit
ggplot(cal_data, aes(x = magnitude_depth_105, y = b_scaled)) +
  geom_point() +                 # Scatter points
  geom_smooth(method = "lm") + # Linear model fit line
  theme_classic()
```

#### B Calls vs. Magnitude at Depth 280

```{r, message = FALSE, warning = FALSE}
# Scatterplot with line of fit
ggplot(cal_data, aes(x = magnitude_depth_280, y = b_scaled)) +
  geom_point() +                 # Scatter points
  geom_smooth(method = "lm") + # Linear model fit line
  theme_classic()
```

#### B Calls vs. Theta at Depth 55

```{r, message = FALSE, warning = FALSE}
# Scatterplot with line of fit
ggplot(cal_data, aes(x = theta_depth_55, y = b_scaled)) +
  geom_point() +                 # Scatter points
  geom_smooth(method = "lm") + # Linear model fit line
  theme_classic()
```

#### B Calls vs. Theta at Depth 105

```{r, message = FALSE, warning = FALSE}
# Scatterplot with line of fit
ggplot(cal_data, aes(x = theta_depth_105, y = b_scaled)) +
  geom_point() +                 # Scatter points
  geom_smooth(method = "lm") + # Linear model fit line
  theme_classic()
```

#### B Calls vs. Theta at Depth 280

```{r, message = FALSE, warning = FALSE}
# Scatterplot with line of fit
ggplot(cal_data, aes(x = theta_depth_280, y = b_scaled)) +
  geom_point() +                 # Scatter points
  geom_smooth(method = "lm") + # Linear model fit line
  theme_classic()
```

### Correlation Plot

```{r}
# Plotting Correlation
# Define expected columns
cols_to_use <- c("est_depth", 
                 "temperature_depth_105", "temperature_depth_280", "temperature_depth_55", 
                 "salinity_depth_105", "salinity_depth_280", "salinity_depth_55", 
                 "meridional_velocity_depth_105", "meridional_velocity_depth_280", "meridional_velocity_depth_55", 
                 "zonal_velocity_depth_105", "zonal_velocity_depth_280", "zonal_velocity_depth_55", 
                 "vertical_velocity_depth_105", "vertical_velocity_depth_280", "vertical_velocity_55", "larvae_count", "lon", "lat", "season")

# Keep only existing columns
existing_cols <- intersect(cols_to_use, colnames(cal_data))
cor_data <- cal_data[, existing_cols, drop = FALSE]

# Compute correlation matrix (use complete.obs to skip missing data)
cor_matrix <- cor(cor_data, use = "complete.obs")

# Correlation plot
ggcorrplot(cor_matrix, 
           lab = TRUE, 
           hc.order = TRUE, 
           lab_size =2,
           tl.cex = 5,
           type = "lower", 
           title = "Correlation Plot")
```

### Missing Data

```{r}
# Count how many missing values there are
sum(is.na(cal_data))

# Visualize missing data
missing_data_plot <- naniar::vis_miss(cal_data)
print(missing_data_plot)
```

### Autocorrelation Detection

```{r}
acf(cal_data$b_scaled)

pacf(cal_data$b_scaled)
```

## Splitting the Data

```{r}
# Split data into training and testing sets with most recent year as our testing set
set.seed(123)
last_year <- max(cal_data$year)
train_data <- cal_data %>% filter(year != last_year)
test_data <- cal_data %>% filter(year == last_year)
```

## Modeling

### Forward Stepwise Selection

```{r}
# Wait for Dr. Baracaldo
```

### Delta-Gamma Model

#### Intercept with `spatial = "on"` and `spatialtemporal = iid`

```{r}
# Create mesh using training data
mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 10)

# Fit spatiotemporal model
fit_spatiotemporal <- sdmTMB(
  b_scaled ~ 1, 
  data = train_data, 
  mesh = mesh_train,
  time = "season",
  family = delta_gamma(link1 = "logit", link2 = "log"), 
  spatial = "on", 
  spatiotemporal = "iid",
  extra_time = unique(test_data$season)
)

# Extract coefficients and sanity check
# tidy(fit_spatiotemporal, conf.int = TRUE)
sanity(fit_spatiotemporal)

# Predictions on training and test data
training_pred <- predict(fit_spatiotemporal, newdata = train_data, type = "response")
training_pred$observed <- train_data$b_scaled

test_pred <- predict(fit_spatiotemporal, newdata = test_data, type = "response")
test_pred$observed <- test_data$b_scaled

# Plotting the predictions
p0 <- ggplot(training_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Train") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

p1 <- ggplot(test_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Test") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

gridExtra::grid.arrange(p0, p1)

# Compute RMSE
rmse_train <- rmse(training_pred$observed, training_pred$est)
rmse_test <- rmse(test_pred$observed, test_pred$est)

# Compute Range
range_train <- range(train_data$b_scaled)
range_test <- range(test_data$b_scaled)

# Compute Standard Deviation
sd_train <- sd(train_data$b_scaled, na.rm = TRUE)
sd_test <- sd(test_data$b_scaled, na.rm = TRUE)

# Assemble the table
summary_table <- data.frame(
  Dataset = c("Training", "Test"),
  RMSE = c(rmse_train, rmse_test),
  Range_Min = c(range_train[1], range_test[1]),
  Range_Max = c(range_train[2], range_test[2]),
  Std_Dev = c(sd_train, sd_test)
)

# Print the summary table
print(summary_table)
```

#### Intercept with `spatial = "off"` and `spatiotemporal = idd`

```{r}
# Create mesh using training data
mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 10)

# Fit spatiotemporal model
fit_spatiotemporal <- sdmTMB(
  b_scaled ~ 1, 
  data = train_data, 
  mesh = mesh_train,
  time = "season",
  family = delta_gamma(link1 = "logit", link2 = "log"), 
  spatial = "off", 
  spatiotemporal = "iid",
  extra_time = unique(test_data$season)
)

# Extract coefficients and sanity check
# tidy(fit_spatiotemporal, conf.int = TRUE)
sanity(fit_spatiotemporal)

# Predictions on training and test data
training_pred <- predict(fit_spatiotemporal, newdata = train_data, type = "response")
training_pred$observed <- train_data$b_scaled

test_pred <- predict(fit_spatiotemporal, newdata = test_data, type = "response")
test_pred$observed <- test_data$b_scaled

# Plotting the predictions
p0 <- ggplot(training_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Train") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

p1 <- ggplot(test_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Test") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

gridExtra::grid.arrange(p0, p1)

# Compute RMSE
rmse_train <- rmse(training_pred$observed, training_pred$est)
rmse_test <- rmse(test_pred$observed, test_pred$est)

# Compute Range
range_train <- range(train_data$b_scaled)
range_test <- range(test_data$b_scaled)

# Compute Standard Deviation
sd_train <- sd(train_data$b_scaled, na.rm = TRUE)
sd_test <- sd(test_data$b_scaled, na.rm = TRUE)

# Assemble the table
summary_table <- data.frame(
  Dataset = c("Training", "Test"),
  RMSE = c(rmse_train, rmse_test),
  Range_Min = c(range_train[1], range_test[1]),
  Range_Max = c(range_train[2], range_test[2]),
  Std_Dev = c(sd_train, sd_test)
)

# Print the summary table
print(summary_table)
```
Intercepts with spatial on vs. spatial off produce the same results for idd. RMSE: 1.6877252.

#### Intercept with `spatial = "on"` and `spatialtemporal = "rw"`

```{r}
# Create mesh using training data
mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 5)

# Fit spatiotemporal model
fit_spatiotemporal <- sdmTMB(
  b_scaled ~ 1, 
  data = train_data, 
  mesh = mesh_train,
  time = "season",
  family = delta_gamma(link1 = "logit", link2 = "log"), 
  spatial = "on", 
  spatiotemporal = "rw",
  extra_time = unique(test_data$season)
)

# Extract coefficients and sanity check
# tidy(fit_spatiotemporal, conf.int = TRUE)
sanity(fit_spatiotemporal)

# Predictions on training and test data
training_pred <- predict(fit_spatiotemporal, newdata = train_data, type = "response")
training_pred$observed <- train_data$b_scaled

test_pred <- predict(fit_spatiotemporal, newdata = test_data, type = "response")
test_pred$observed <- test_data$b_scaled

# Plotting the predictions
p0 <- ggplot(training_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Train") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

p1 <- ggplot(test_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Test") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

gridExtra::grid.arrange(p0, p1)

# Compute RMSE
rmse_train <- rmse(training_pred$observed, training_pred$est)
rmse_test <- rmse(test_pred$observed, test_pred$est)

# Compute Range
range_train <- range(train_data$b_scaled)
range_test <- range(test_data$b_scaled)

# Compute Standard Deviation
sd_train <- sd(train_data$b_scaled, na.rm = TRUE)
sd_test <- sd(test_data$b_scaled, na.rm = TRUE)

# Assemble the table
summary_table <- data.frame(
  Dataset = c("Training", "Test"),
  RMSE = c(rmse_train, rmse_test),
  Range_Min = c(range_train[1], range_test[1]),
  Range_Max = c(range_train[2], range_test[2]),
  Std_Dev = c(sd_train, sd_test)
)

# Print the summary table
print(summary_table)
```
Performed well. Test RMSE: 1.6790743 and test st dev: 1.669873 for mesh cutoff = 10.
Test RMSE: 1.3885117 and test st dev: 1.482606 for mesh cutoff = 5.

#### Intercept with `spatial = "off"` and `spatialtemporal = "rw"`

```{r}
# Create mesh using training data
mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 10)

# Fit spatiotemporal model
fit_spatiotemporal <- sdmTMB(
  b_scaled ~ 1, 
  data = train_data, 
  mesh = mesh_train,
  time = "season",
  family = delta_gamma(link1 = "logit", link2 = "log"), 
  spatial = "off", 
  spatiotemporal = "rw",
  extra_time = unique(test_data$season)
)

# Extract coefficients and sanity check
# tidy(fit_spatiotemporal, conf.int = TRUE)
sanity(fit_spatiotemporal)

# Predictions on training and test data
training_pred <- predict(fit_spatiotemporal, newdata = train_data, type = "response")
training_pred$observed <- train_data$b_scaled

test_pred <- predict(fit_spatiotemporal, newdata = test_data, type = "response")
test_pred$observed <- test_data$b_scaled

# Plotting the predictions
p0 <- ggplot(training_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Train") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

p1 <- ggplot(test_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Test") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

gridExtra::grid.arrange(p0, p1)

# Compute RMSE
rmse_train <- rmse(training_pred$observed, training_pred$est)
rmse_test <- rmse(test_pred$observed, test_pred$est)

# Compute Range
range_train <- range(train_data$b_scaled)
range_test <- range(test_data$b_scaled)

# Compute Standard Deviation
sd_train <- sd(train_data$b_scaled, na.rm = TRUE)
sd_test <- sd(test_data$b_scaled, na.rm = TRUE)

# Assemble the table
summary_table <- data.frame(
  Dataset = c("Training", "Test"),
  RMSE = c(rmse_train, rmse_test),
  Range_Min = c(range_train[1], range_test[1]),
  Range_Max = c(range_train[2], range_test[2]),
  Std_Dev = c(sd_train, sd_test)
)

# Print the summary table
print(summary_table)
```
Spatial on performs better than spatial off with rw. RMSE: 1.7099783.

#### Intercept with `spatial = "on"` and `spatiotemporal = "ar1"`

```{r}
# Create mesh using training data
mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 10)

# Fit spatiotemporal model
fit_spatiotemporal <- sdmTMB(
  b_scaled ~ 1, 
  data = train_data, 
  mesh = mesh_train,
  time = "season",
  family = delta_gamma(link1 = "logit", link2 = "log"), 
  spatial = "on", 
  spatiotemporal = "ar1",
  extra_time = unique(test_data$season)
)

# Extract coefficients and sanity check
# tidy(fit_spatiotemporal, conf.int = TRUE)
sanity(fit_spatiotemporal)

# Predictions on training and test data
training_pred <- predict(fit_spatiotemporal, newdata = train_data, type = "response")
training_pred$observed <- train_data$b_scaled

test_pred <- predict(fit_spatiotemporal, newdata = test_data, type = "response")
test_pred$observed <- test_data$b_scaled

# Plotting the predictions
p0 <- ggplot(training_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Train") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

p1 <- ggplot(test_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Test") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

gridExtra::grid.arrange(p0, p1)

# Compute RMSE
rmse_train <- rmse(training_pred$observed, training_pred$est)
rmse_test <- rmse(test_pred$observed, test_pred$est)

# Compute Range
range_train <- range(train_data$b_scaled)
range_test <- range(test_data$b_scaled)

# Compute Standard Deviation
sd_train <- sd(train_data$b_scaled, na.rm = TRUE)
sd_test <- sd(test_data$b_scaled, na.rm = TRUE)

# Assemble the table
summary_table <- data.frame(
  Dataset = c("Training", "Test"),
  RMSE = c(rmse_train, rmse_test),
  Range_Min = c(range_train[1], range_test[1]),
  Range_Max = c(range_train[2], range_test[2]),
  Std_Dev = c(sd_train, sd_test)
)

# Print the summary table
print(summary_table)
```
Causes an error.

#### Intercept with `spatial = "off"` and `spatiotemporal = "ar1"`

```{r}
# Create mesh using training data
mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 10)

# Fit spatiotemporal model
fit_spatiotemporal <- sdmTMB(
  b_scaled ~ 1, 
  data = train_data, 
  mesh = mesh_train,
  time = "season",
  family = delta_gamma(link1 = "logit", link2 = "log"), 
  spatial = "off", 
  spatiotemporal = "ar1",
  extra_time = unique(test_data$season)
)

# Extract coefficients and sanity check
# tidy(fit_spatiotemporal, conf.int = TRUE)
sanity(fit_spatiotemporal)

# Predictions on training and test data
training_pred <- predict(fit_spatiotemporal, newdata = train_data, type = "response")
training_pred$observed <- train_data$b_scaled

test_pred <- predict(fit_spatiotemporal, newdata = test_data, type = "response")
test_pred$observed <- test_data$b_scaled

# Plotting the predictions
p0 <- ggplot(training_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Train") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

p1 <- ggplot(test_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Test") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

gridExtra::grid.arrange(p0, p1)

# Compute RMSE
rmse_train <- rmse(training_pred$observed, training_pred$est)
rmse_test <- rmse(test_pred$observed, test_pred$est)

# Compute Range
range_train <- range(train_data$b_scaled)
range_test <- range(test_data$b_scaled)

# Compute Standard Deviation
sd_train <- sd(train_data$b_scaled, na.rm = TRUE)
sd_test <- sd(test_data$b_scaled, na.rm = TRUE)

# Assemble the table
summary_table <- data.frame(
  Dataset = c("Training", "Test"),
  RMSE = c(rmse_train, rmse_test),
  Range_Min = c(range_train[1], range_test[1]),
  Range_Max = c(range_train[2], range_test[2]),
  Std_Dev = c(sd_train, sd_test)
)

# Print the summary table
print(summary_table)
```
Performs the same as the iid model. RMSE: 1.6877254.

#### Environmental Predictors with `spatial = "on"` and `spatiotemporal = "iid"`

```{r}
# Create mesh using training data
mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 10)

# Fit spatiotemporal model
fit_spatiotemporal <- sdmTMB(
  b_scaled ~ temperature_depth_105_std + temperature_depth_55_std + temperature_depth_280_std + salinity_depth_55_std + salinity_depth_105_std + magnitude_depth_55_std + magnitude_depth_105_std + theta_depth_55_std + theta_depth_105_std,  
  data = train_data, 
  mesh = mesh_train,
  time = "season",
  family = delta_gamma(link1 = "logit", link2 = "log"), 
  spatial = "on", 
  spatiotemporal = "iid",
  extra_time = unique(test_data$season)
)

# Extract coefficients and sanity check
#tidy(fit_spatiotemporal, conf.int = TRUE)
sanity(fit_spatiotemporal)

# Predictions on training and test data
training_pred <- predict(fit_spatiotemporal, newdata = train_data, type = "response")
training_pred$observed <- train_data$b_scaled

test_pred <- predict(fit_spatiotemporal, newdata = test_data, type = "response")
test_pred$observed <- test_data$b_scaled

# Plotting the predictions
p0 <- ggplot(training_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Train") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

p1 <- ggplot(test_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Test") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

gridExtra::grid.arrange(p0, p1)

# Compute RMSE
rmse_train <- rmse(training_pred$observed, training_pred$est)
rmse_test <- rmse(test_pred$observed, test_pred$est)

# Compute Range
range_train <- range(train_data$b_scaled, na.rm = TRUE)
range_test <- range(test_data$b_scaled, na.rm = TRUE)

# Compute Standard Deviation
sd_train <- sd(train_data$b_scaled, na.rm = TRUE)
sd_test <- sd(test_data$b_scaled, na.rm = TRUE)

# Assemble the table
summary_table <- data.frame(
  Dataset = c("Training", "Test"),
  RMSE = c(rmse_train, rmse_test),
  Range_Min = c(range_train[1], range_test[1]),
  Range_Max = c(range_train[2], range_test[2]),
  Std_Dev = c(sd_train, sd_test)
)

# Print the summary table
print(summary_table)
```
Performs the same as the other models. RMSE: 1.6877254.

#### Environmental Predictors with `spatial = "off"` and `spatiotemporal = "idd"`

```{r}
# Create mesh using training data
mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 10)

# Fit spatiotemporal model
fit_spatiotemporal <- sdmTMB(
  b_scaled ~ s(temperature_depth_105_std) + s(temperature_depth_55_std) + s(temperature_depth_280_std) + s(salinity_depth_55_std) + s(salinity_depth_105_std) + s(magnitude_depth_55_std) + s(magnitude_depth_105_std) + s(theta_depth_55_std) + s(theta_depth_105_std),  
  data = train_data, 
  mesh = mesh_train,
  time = "season",
  family = delta_gamma(link1 = "logit", link2 = "log"), 
  spatial = "off", 
  spatiotemporal = "iid",
  extra_time = unique(test_data$season)
)

# Extract coefficients and sanity check
tidy(fit_spatiotemporal, conf.int = TRUE)
sanity(fit_spatiotemporal)

# Predictions on training and test data
training_pred <- predict(fit_spatiotemporal, newdata = train_data, type = "response")
training_pred$observed <- train_data$b_scaled

test_pred <- predict(fit_spatiotemporal, newdata = test_data, type = "response")
test_pred$observed <- test_data$b_scaled

# Plotting the predictions
p0 <- ggplot(training_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Train") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

p1 <- ggplot(test_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Test") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

gridExtra::grid.arrange(p0, p1)

# Compute RMSE
rmse_train <- rmse(training_pred$observed, training_pred$est)
rmse_test <- rmse(test_pred$observed, test_pred$est)

# Compute Range
range_train <- range(train_data$b_scaled, na.rm = TRUE)
range_test <- range(test_data$b_scaled, na.rm = TRUE)

# Compute Standard Deviation
sd_train <- sd(train_data$b_scaled, na.rm = TRUE)
sd_test <- sd(test_data$b_scaled, na.rm = TRUE)

# Assemble the table
summary_table <- data.frame(
  Dataset = c("Training", "Test"),
  RMSE = c(rmse_train, rmse_test),
  Range_Min = c(range_train[1], range_test[1]),
  Range_Max = c(range_train[2], range_test[2]),
  Std_Dev = c(sd_train, sd_test)
)

# Print the summary table
print(summary_table)
```
RMSE: 1.678275. iid unchanged by splines.

#### Environmental Predictors with `spatial = "on"` and `spatiotemporal = "rw"`

```{r}
# Create mesh using training data
mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 5)

# Fit spatiotemporal model
fit_spatiotemporal <- sdmTMB(
  b_scaled ~ s(temperature_depth_105_std) + s(temperature_depth_55_std) + s(temperature_depth_280_std) + s(salinity_depth_55_std) + s(salinity_depth_105_std) + s(magnitude_depth_55_std) + s(magnitude_depth_105_std) + s(theta_depth_55_std) + s(theta_depth_105_std),  
  data = train_data, 
  mesh = mesh_train,
  time = "season",
  family = delta_gamma(link1 = "logit", link2 = "log"),
  spatial = "on", 
  spatiotemporal = "rw",
  extra_time = unique(test_data$season)
)

# Extract coefficients and sanity check
tidy(fit_spatiotemporal, conf.int = TRUE)
sanity(fit_spatiotemporal)

# Predictions on training and test data
training_pred <- predict(fit_spatiotemporal, newdata = train_data, type = "response")
training_pred$observed <- train_data$b_scaled

test_pred <- predict(fit_spatiotemporal, newdata = test_data, type = "response")
test_pred$observed <- test_data$b_scaled

# Plotting the predictions
p0 <- ggplot(training_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Train") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

p1 <- ggplot(test_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Test") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

gridExtra::grid.arrange(p0, p1)

# Compute RMSE
rmse_train <- rmse(training_pred$observed, training_pred$est)
rmse_test <- rmse(test_pred$observed, test_pred$est)

# Compute Range
range_train <- range(train_data$b_scaled, na.rm = TRUE)
range_test <- range(test_data$b_scaled, na.rm = TRUE)

# Compute Standard Deviation
sd_train <- sd(train_data$b_scaled, na.rm = TRUE)
sd_test <- sd(test_data$b_scaled, na.rm = TRUE)

# Assemble the table
summary_table <- data.frame(
  Dataset = c("Training", "Test"),
  RMSE = c(rmse_train, rmse_test),
  Range_Min = c(range_train[1], range_test[1]),
  Range_Max = c(range_train[2], range_test[2]),
  Std_Dev = c(sd_train, sd_test)
)

# Print the summary table
print(summary_table)
```
Spatial on and rw is performing the best. Test RMSE: 1.491210 and test st dev: 1.669873 for mesh cutoff = 10.
Test RMSE: 1.491210 and test st dev: 1.669873 for mesh cutoff = 5. Unchanged by mesh.
Test RMSE: 1.459567	and st dev: 1.669873 with splines.

#### Environmental Predictors with `spatial = "off"` and `spatiotemporal = "rw"`

```{r}
# Create mesh using training data
mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 10)

# Fit spatiotemporal model
fit_spatiotemporal <- sdmTMB(
  b_scaled ~ temperature_depth_105_std + temperature_depth_55_std + temperature_depth_280_std + salinity_depth_55_std + salinity_depth_105_std + magnitude_depth_55_std + magnitude_depth_105_std + theta_depth_55_std + theta_depth_105_std,  
  data = train_data, 
  mesh = mesh_train,
  time = "season",
  family = delta_gamma(link1 = "logit", link2 = "log"),
  spatial = "off", 
  spatiotemporal = "rw",
  extra_time = unique(test_data$season)
)

# Extract coefficients and sanity check
# tidy(fit_spatiotemporal, conf.int = TRUE)
sanity(fit_spatiotemporal)

# Predictions on training and test data
training_pred <- predict(fit_spatiotemporal, newdata = train_data, type = "response")
training_pred$observed <- train_data$b_scaled

test_pred <- predict(fit_spatiotemporal, newdata = test_data, type = "response")
test_pred$observed <- test_data$b_scaled

# Plotting the predictions
p0 <- ggplot(training_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Train") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

p1 <- ggplot(test_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Test") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

gridExtra::grid.arrange(p0, p1)

# Compute RMSE
rmse_train <- rmse(training_pred$observed, training_pred$est)
rmse_test <- rmse(test_pred$observed, test_pred$est)

# Compute Range
range_train <- range(train_data$b_scaled, na.rm = TRUE)
range_test <- range(test_data$b_scaled, na.rm = TRUE)

# Compute Standard Deviation
sd_train <- sd(train_data$b_scaled, na.rm = TRUE)
sd_test <- sd(test_data$b_scaled, na.rm = TRUE)

# Assemble the table
summary_table <- data.frame(
  Dataset = c("Training", "Test"),
  RMSE = c(rmse_train, rmse_test),
  Range_Min = c(range_train[1], range_test[1]),
  Range_Max = c(range_train[2], range_test[2]),
  Std_Dev = c(sd_train, sd_test)
)

# Print the summary table
print(summary_table)
```
Performed okay. RMSE: 1.500728.

#### Environmental Predictors with `spatial = "on"` and `spatiotemporal = "ar1"`

```{r}
# Create mesh using training data
mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 10)

# Fit spatiotemporal model
fit_spatiotemporal <- sdmTMB(
  b_scaled ~ temperature_depth_105_std + temperature_depth_55_std + temperature_depth_280_std + salinity_depth_55_std + salinity_depth_105_std + magnitude_depth_55_std + magnitude_depth_105_std + theta_depth_55_std + theta_depth_105_std,  
  data = train_data, 
  mesh = mesh_train,
  time = "season",
  family = delta_gamma(link1 = "logit", link2 = "log"),
  spatial = "on", 
  spatiotemporal = "ar1",
  extra_time = unique(test_data$season)
)

# Extract coefficients and sanity check
# tidy(fit_spatiotemporal, conf.int = TRUE)
sanity(fit_spatiotemporal)

# Predictions on training and test data
training_pred <- predict(fit_spatiotemporal, newdata = train_data, type = "response")
training_pred$observed <- train_data$b_scaled

test_pred <- predict(fit_spatiotemporal, newdata = test_data, type = "response")
test_pred$observed <- test_data$b_scaled

# Plotting the predictions
p0 <- ggplot(training_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Train") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

p1 <- ggplot(test_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Test") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

gridExtra::grid.arrange(p0, p1)

# Compute RMSE
rmse_train <- rmse(training_pred$observed, training_pred$est)
rmse_test <- rmse(test_pred$observed, test_pred$est)

# Compute Range
range_train <- range(train_data$b_scaled, na.rm = TRUE)
range_test <- range(test_data$b_scaled, na.rm = TRUE)

# Compute Standard Deviation
sd_train <- sd(train_data$b_scaled, na.rm = TRUE)
sd_test <- sd(test_data$b_scaled, na.rm = TRUE)

# Assemble the table
summary_table <- data.frame(
  Dataset = c("Training", "Test"),
  RMSE = c(rmse_train, rmse_test),
  Range_Min = c(range_train[1], range_test[1]),
  Range_Max = c(range_train[2], range_test[2]),
  Std_Dev = c(sd_train, sd_test)
)

# Print the summary table
print(summary_table)
```
Wouldn't stop running.

#### Environmental Predictors with `spatial = "off"` and `spatiotemporal = "ar1"`

```{r}
# Create mesh using training data
mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 10)

# Fit spatiotemporal model
fit_spatiotemporal <- sdmTMB(
  b_scaled ~ temperature_depth_105_std + temperature_depth_55_std + temperature_depth_280_std + salinity_depth_55_std + salinity_depth_105_std + magnitude_depth_55_std + magnitude_depth_105_std + theta_depth_55_std + theta_depth_105_std,  
  data = train_data, 
  mesh = mesh_train,
  time = "season",
  family = delta_gamma(link1 = "logit", link2 = "log"),
  spatial = "off", 
  spatiotemporal = "ar1",
  extra_time = unique(test_data$season)
)

# Extract coefficients and sanity check
# tidy(fit_spatiotemporal, conf.int = TRUE)
sanity(fit_spatiotemporal)

# Predictions on training and test data
training_pred <- predict(fit_spatiotemporal, newdata = train_data, type = "response")
training_pred$observed <- train_data$b_scaled

test_pred <- predict(fit_spatiotemporal, newdata = test_data, type = "response")
test_pred$observed <- test_data$b_scaled

# Plotting the predictions
p0 <- ggplot(training_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Train") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

p1 <- ggplot(test_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Test") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

gridExtra::grid.arrange(p0, p1)

# Compute RMSE
rmse_train <- rmse(training_pred$observed, training_pred$est)
rmse_test <- rmse(test_pred$observed, test_pred$est)

# Compute Range
range_train <- range(train_data$b_scaled, na.rm = TRUE)
range_test <- range(test_data$b_scaled, na.rm = TRUE)

# Compute Standard Deviation
sd_train <- sd(train_data$b_scaled, na.rm = TRUE)
sd_test <- sd(test_data$b_scaled, na.rm = TRUE)

# Assemble the table
summary_table <- data.frame(
  Dataset = c("Training", "Test"),
  RMSE = c(rmse_train, rmse_test),
  Range_Min = c(range_train[1], range_test[1]),
  Range_Max = c(range_train[2], range_test[2]),
  Std_Dev = c(sd_train, sd_test)
)

# Print the summary table
print(summary_table)
```
Caused an error.
