# Justin Individual Report 


## Summary of Contributions 

Throughout the project, I took on the task of adding in data, cleaning data, modeling blue whale calls, and a majority of my efforts in creating visualizations. 



## Cleaning data 

In the beginning, Michaela Alksne provided us with CalCOFI acoustic data. We decided that it would be best if we could combine them and clean the column names. I wrote a script that would read in all the merged acoustic data and clean the column names for future 


```{r eval = FALSE} 

# View the merged data
view(merged_CalCOFI)

colnames(merged_CalCOFI)[4] = "Station_Key"
# Iterate through column names and remove duplicate columns 
merged_CalCOFI <- merged_CalCOFI[, !duplicated(colnames(merged_CalCOFI))]

# View the merged data
view(merged_CalCOFI)

# Change column names to be snake case
colnames(merged_CalCOFI) <- tolower(colnames(merged_CalCOFI))

# If column names have spaces, period or commas, replace with underscores
colnames(merged_CalCOFI) <- gsub(" ", "_", colnames(merged_CalCOFI))
colnames(merged_CalCOFI) <- gsub("\\.", "_", colnames(merged_CalCOFI))
colnames(merged_CalCOFI) <- gsub(",", "_", colnames(merged_CalCOFI))

# View the merged data
view(merged_CalCOFI)
```




## Adding Euphasia and Larvae Data 




## Acoustic Data Visuaslization

## Blue Whale Calls


A part of my work was to also work with the A, B, D calls to model and predict Blue Whale Calls on the southern coast. I first prepared the data by assigning the seasons correctly: Winter would have December, January, February. After which, I mutated the A, B, D calls and added them together to get the total whale calls. This way, I was able to get my actual response variable which was the total collective amount of blue whale calls at any given moment 


```{r eval = FALSE}
raw_data <- read.csv("../../data/Merged_CalCOFI/STD_CalCOFI_final")
# Create a new `seasonal_category` column
prep_data <- raw_data %>% 
  mutate(seasonal_category = case_when(
    month %in% c(12, 1, 2)  ~ "Winter",
    month %in% c(3, 4, 5)   ~ "Spring",
    month %in% c(6, 7, 8)   ~ "Summer",
    month %in% c(9, 10, 11) ~ "Fall"
  ))

# Removing the first two columns
prep_data <- prep_data[, -c(1, 2)]
# Create fin whale response column
prep_data <- prep_data %>%
  mutate(blue_resp = a_scaled + b_scaled + d_scaled)

```

I then attempted to do exploratory data analysis on the data in order to see if we can visually detect any trends between the predictors and the response variables. The graph would plot each individual call with a different color for better visibility of the distribution of each call 

```{r eval = FALSE}
library(ggExtra)

predictors <- c(
  "lon", "lat", "date", "season", "est_depth", "larvae_count",
  "temperature_depth_55", "temperature_depth_105", "temperature_depth_280",
  "pressure_depth_55", "pressure_depth_105", "pressure_depth_280",
  "salinity_depth_55", "salinity_depth_105", "salinity_depth_280",
  "magnitude_depth_55", "magnitude_depth_105", "magnitude_depth_280",
  "theta_depth_55", "theta_depth_105", "theta_depth_280"
)




library(ggExtra)
library(gridExtra)

# Prepare data
long_cal_data <- cal_data %>%
  select(one_of(predictors), a_scaled, b_scaled, d_scaled, blue_resp) %>%
  pivot_longer(cols = c(a_scaled, b_scaled, d_scaled),
               names_to = "source", values_to = "contribution") %>%
  filter(!is.na(contribution), contribution > 5)

# List to hold ggMarginal plots (ggExtraPlot objects)
plot_list <- list()

for (var in predictors) {
  clean_data <- long_cal_data[!is.na(long_cal_data[[var]]) & !is.na(long_cal_data$blue_resp), ]
  
  base_plot <- ggplot(clean_data, aes_string(x = var, y = "contribution", color = "source")) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "lm", se = FALSE) +
    theme_classic() +
    ggtitle(paste(var))
  
  marginal_plot <- ggMarginal(base_plot, type = "histogram", groupColour = TRUE, groupFill = TRUE)
  
  plot_list[[var]] <- marginal_plot  # Store the full ggExtraPlot object
}
# 
# # Plot the first 6 in a 3x2 grid
# do.call(grid.arrange, c(plot_list[1:4], ncol = 2))

# Optional: Loop through all in chunks
chunks <- split(plot_list, ceiling(seq_along(plot_list) / 4))
for (chunk in chunks) {
  do.call(grid.arrange, c(chunk, ncol = 2))
}
```

Throughout the graphs, I did realize that the calls were centered more towards a middle value rather than random. 



After the exploratory data analysis, I then tried to make a 


```{r eval = FALSE}
set.seed(123)
last_year <- c(2023)

# Changing dataset to only have present response data
cal_data <- prep_data %>% 
  filter(!is.na(blue_resp))

train_data <- cal_data %>% filter(!year %in% last_year)
test_data  <- cal_data %>% filter(year %in% last_year)
# DELTA GAMMA
mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 10)
# Fit spatiotemporal model
fit <- sdmTMB(
  blue_resp ~ lat + lon + temperature_depth_55_std + salinity_depth_55_std,
  data = train_data,
  mesh = mesh_train,
  time = "season",
  family = delta_gamma(link1 = "logit", link2 = "log"),
  spatial = "on",
  spatiotemporal = "iid",
  extra_time = seq(2023.00, 2023.75, by = 0.25)
)

blue_plots <- generate_seasonal_prediction_plots(fit, year = "2023", test_data = test_data, cap = 2, resp = "blue_resp")
wrap_plots(blue_plots, ncol = 2, nrow = 2, guides = "collect") +
  plot_annotation(
    title = "Seasonal Blue Whale Density Predictions (2023)",
    theme = theme(plot.title = element_text(hjust = 0.5, size = 16))
  )

```

## Whale Call Model Visualizaition
