---
title: "D_call_final_report"
author: "Peter Xiong"
date: "2025-05-15"
output: html_document
---

## Load Packages and Data

```{r, message = F, warning = F}
# Load in the libraries
library(tidyverse)
library(sdmTMB)
library(ggplot2)
library(janitor)
library(sp)
library(readr)
library(ggcorrplot)
library(Metrics)
library(naniar)
library(tibble)
library(rsample)
library(caret)
library(purrr)
library(tune)
library(ggExtra)
library(gridExtra)

# Libraries for plotting
library(sf)
library(rnaturalearth)
library(viridis)
library(patchwork)

# Read in the data
cal_data <- read.csv("D:/Desktop/PSTAT197B/scripps/capstone-scripps/data/Merged_CalCOFI/merged_std_data_05_13_25.csv")
```

## Data Cleaning

```{r}
# Removing the first two columns
cal_data <- cal_data[, -c(1, 2)]

# Changing dataset to only have present response data
cal_data <- cal_data %>% 
  filter(!is.na(d_scaled)) %>%
  mutate(blue_resp = a_scaled + b_scaled + d_scaled)
```

## Exploratory Data Analysis

### Response vs Predictors

Before modeling, I want to explore the relationship between the response variable and predictors.

First, we need to filter out all zero values:

```{r}
d_filtered <- cal_data %>% filter(d_scaled!= 0)
```

#### D calls vs. Lon&Lat

```{r, message = FALSE}
p1 <- ggplot(d_filtered, aes(x = lon, y = d_scaled)) +
  geom_point() +
  theme_classic()

p1_marginal <- ggMarginal(p1, type = "histogram")

p2 <- ggplot(d_filtered, aes(x = lat, y = d_scaled)) +
  geom_point() +
  theme_classic()

p2_marginal <- ggMarginal(p2, type = "histogram")

## Arrange them in a single row
gridExtra::grid.arrange(p1_marginal, p2_marginal)
```

D-call counts are almost always low or zero and only occasionally spike to high values. Those spikes do not concentrate at a single latitude or longitude, so neither coordinate shows a clear linear trend with call intensity.

#### D calls vs. Season

```{r, message=FALSE}
p3 <- ggplot(d_filtered, aes(x = season_decimal, y = d_scaled)) +
  geom_point() +
  theme_classic()

ggMarginal(p3, type = "histogram")
```

From the plot, we can see that most D calls were detected in spring and summer. It means that blue whales are more active in spring and summer.

#### D calls vs. Temperature

```{r, message=FALSE}
p4 <- ggplot(d_filtered, aes(x = temperature_depth_55, y = d_scaled)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(title = "D calls vs. Temperature at Depth 55m")+
  theme_classic()

p5 <- ggplot(d_filtered, aes(x = temperature_depth_105, y = d_scaled)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(title = "D calls vs. Temperature at Depth 105m")+
  theme_classic()

p6 <- ggplot(d_filtered, aes(x = temperature_depth_280, y = d_scaled)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(title = "D calls vs. Temperature at Depth 280m")+
  theme_classic()

## Arrange them in a single row
gridExtra::grid.arrange(p4, p5, p6)
```

Across 55 m, 105 m, and 280 m, water temperature spans roughly 12 – 20 °C and the three depth profiles look almost identical—reflecting the strong inter-depth correlations seen in the matrix. In every panel most D-call counts stay below 5, with a heavy tail of zeros and only two or three extreme outliers (> 30) that are scattered across the full temperature range.

#### D calls vs. Pressure

```{r, message=FALSE}
p7 <- ggplot(d_filtered, aes(x = pressure_depth_55, y = d_scaled)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(title = "D calls vs. Pressure at Depth 55m")+
  theme_classic()

p8 <- ggplot(d_filtered, aes(x = pressure_depth_105, y = d_scaled)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(title = "D calls vs. Pressure at Depth 105m")+
  theme_classic()

p9 <- ggplot(d_filtered, aes(x = pressure_depth_280, y = d_scaled)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(title = "D calls vs. Pressure at Depth 280m")+
  theme_classic()

## Arrange them in a single row
gridExtra::grid.arrange(p7, p8, p9)
```

Across the 55 m, 105 m, and 280 m layers, the smoothed blue curves remain essentially flat, with confidence bands overlapping zero throughout the observed pressure range. Scattered points—including a few high-count outliers—occur at various pressure values without a clear pattern. The similarity of the three panels reinforces that depth-specific pressure offers little additional information for explaining D-call intensity. In short, ambient pressure shows no meaningful relationship with blue-whale D-call activity in this data set.

#### D calls vs. Salinity

```{r, message=FALSE}
s1 <- ggplot(d_filtered, aes(x = salinity_depth_55, y = d_scaled)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(title = "D calls vs. Salinity at Depth 55m")+
  theme_classic()

s2 <- ggplot(d_filtered, aes(x = salinity_depth_105, y = d_scaled)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(title = "D calls vs. Salinity at Depth 105m")+
  theme_classic()

s3 <- ggplot(d_filtered, aes(x = salinity_depth_280, y = d_scaled)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(title = "D calls vs. Salinity at Depth 280m")+
  theme_classic()

## Arrange them in a single row
gridExtra::grid.arrange(s1, s2, s3)
```

For salinity values ranging ≈ 33.2–33.7 PSU, the smoothed trend lines at 55 m, 105 m, and 280 m are virtually flat, with confidence bands overlapping zero along the entire axis. Most D-call counts remain below 5, and the few larger outliers appear at assorted salinities. Because all three depths yield the same, near-zero slope, salinity—like temperature and pressure—shows no discernible influence on D-call intensity in this dataset, and separate depth layers add no extra explanatory power.

#### D calls vs. Magnitude

```{r, message=FALSE}
m1 <- ggplot(d_filtered, aes(x = magnitude_depth_55, y = d_scaled)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(title = "D calls vs. Magnitude at Depth 55m")+
  theme_classic()

m2 <- ggplot(d_filtered, aes(x = magnitude_depth_105, y = d_scaled)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(title = "D calls vs. Magnitude at Depth 105m")+
  theme_classic()

m3 <- ggplot(d_filtered, aes(x = magnitude_depth_280, y = d_scaled)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(title = "D calls vs. Magnitude at Depth 280m")+
  theme_classic()

## Arrange them in a single row
gridExtra::grid.arrange(m1, m2, m3)
```

Current magnitude (0–0.35 m s⁻¹) shows virtually no association with D-call intensity at 55 m, 105 m, or 280 m. Smoothed lines are flat and confidence bands straddle zero throughout, while most detections remain under 5 regardless of flow speed. A few high outliers appear but do not cluster at any particular magnitude. Thus, current speed—like the other environmental variables examined—adds little explanatory power for predicting D-call counts, and additional depth layers do not change that conclusion.

#### D calls vs. Theta

```{r, message=FALSE}
t1 <- ggplot(d_filtered, aes(x = theta_depth_55, y = d_scaled)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(title = "D calls vs. Theta at Depth 55m")+
  theme_classic()

t2 <- ggplot(d_filtered, aes(x = theta_depth_105, y = d_scaled)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(title = "D calls vs. Theta at Depth 105m")+
  theme_classic()

t3 <- ggplot(d_filtered, aes(x = theta_depth_280, y = d_scaled)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(title = "D calls vs. Theta at Depth 280m")+
  theme_classic()

## Arrange them in a single row
gridExtra::grid.arrange(t1, t2, t3)
```

The direction of flow (θ, scaled between –π and +π) shows the same story at 55 m, 105 m, and 280 m: the fitted trend is essentially flat with confidence bands that straddle zero across all directions. Most detections cluster near zero counts, and the handful of large outliers are scattered at various θ values rather than concentrated in any quadrant. Because the three depth profiles are nearly identical, current direction provides little independent information for predicting D-call intensity, confirming that directional flow is not a key driver in this dataset.

### Correlation Plot

```{r}
# Plotting Correlation
# Define expected columns
cols_to_use <- c("est_depth", 
                 "temperature_depth_105", "temperature_depth_280", "temperature_depth_55", 
                 "salinity_depth_105", "salinity_depth_280", "salinity_depth_55", 
                 "meridional_velocity_depth_105", "meridional_velocity_depth_280", "meridional_velocity_depth_55", 
                 "zonal_velocity_depth_105", "zonal_velocity_depth_280", "zonal_velocity_depth_55", 
                 "vertical_velocity_depth_105", "vertical_velocity_depth_280", "vertical_velocity_55", "larvae_count", "lon", "lat", "season")

# Keep only existing columns
existing_cols <- intersect(cols_to_use, colnames(cal_data))
cor_data <- cal_data[, existing_cols, drop = FALSE]

# Compute correlation matrix (use complete.obs to skip missing data)
cor_matrix <- cor(cor_data, use = "complete.obs")

# Correlation plot
ggcorrplot(cor_matrix, 
           lab = TRUE, 
           hc.order = TRUE, 
           lab_size =2,
           tl.cex = 5,
           type = "lower", 
           title = "Correlation Plot")
```

The correlation matrix reveals two clear patterns. First, depth-stratified versions of the same variable (e.g., temperature at 55 m, 105 m, and 280 m) are almost perfectly correlated ( r ≈ 0.9 – 1.0 ), while correlations between different variable types (temperature vs salinity, temperature vs velocity, etc.) hover near zero. Second, the velocity components and spatial coordinates (latitude, longitude) show only weak to moderate associations with any environmental variable.

For modelling, this means that including multiple depths of a highly correlated variable will add redundancy and inflates multicollinearity, potentially causing singular fits or unstable coefficients. A practical remedy is to keep just one representative depth per variable (or combine depths through PCA/regularisation), while retaining spatial terms that contribute largely independent information.

### Missing Data

We need to make sure that there is no missing data.

```{r}
# Count how many missing values there are
sum(is.na(cal_data))

# Visualize missing data
missing_data_plot <- naniar::vis_miss(cal_data)
print(missing_data_plot)
```


## Splitting the Data

```{r}
# Split data into training and testing sets with most recent year as our testing set
set.seed(123)
last_year <- max(cal_data$year)
train_data <- cal_data %>% filter(year != last_year)
test_data <- cal_data %>% filter(year == last_year)
```

## Prediction Plot

#### Function

To determine the performance of models, we want to make a prediction plot to compare the predicted whale existence with the real one. Thus, we need to write a function for plotting.

```{r}
generate_seasonal_prediction_plots <- function(fit, year, test_data, cap, resp) {
  
  plots <- list()
  coast <- rnaturalearth::ne_coastline(scale = "medium", returnclass = "sf")
  choose_folder <- function(variable, year) {
    base_path <- "D:/Desktop/PSTAT197B/scripps/capstone-scripps/data/Specificdate"
    if (variable == "temperature") {
      return(file.path(base_path, "Temperature", "temperature_avg_output"))
    }
    if (variable == "salinity") {
      return(file.path(base_path, "Salinity", "salinity_avg_output"))
    }
    if (variable == "pressure") {
      return(file.path(base_path, "Pressure", "pressure_avg_output"))
    }
    if (variable == "mvel") {
      return(file.path(base_path, "Meridional_velocity", "meridional_velocity_avg_output"))
    }
    if (variable == "zvel") {
      return(file.path(base_path, "Zonal_velocity", "zonal_velocity_avg_output"))
    }
  }

  pivot_grid <- function(df, var_name) {
    df %>%
      rename(lon = 1) %>%
      pivot_longer(cols = starts_with("Lat_"), names_to = "lat", values_to = var_name) %>%
      mutate(
        lat = as.numeric(str_remove(lat, "Lat_")),
        lon = as.numeric(str_remove(lon, "Lon_")) - 360,
        !!var_name := if (!str_detect(var_name, "meridional_velocity|zonal_velocity"))
          scale(.data[[var_name]])[, 1] else .data[[var_name]]
      )
  }

  get_season <- function(month) {
    case_when(
      month %in% c(12, 1, 2) ~ 0,
      month %in% 3:5 ~ 0.25,
      month %in% 6:8 ~ 0.5,
      month %in% 9:11 ~ 0.75
    )
  }

  for (season_val in c(0, 0.25, 0.5, 0.75)) {
    # issue in that the 12 is 12 of that year -> need to hardcode 12 of previous year 
    season_months <- list(`0` = c(12, 1, 2), `0.25` = 3:5, `0.5` = 6:8, `0.75` = 9:11)[[as.character(season_val)]]
    season_data <- list()

    for (month in season_months) {
      date <- sprintf("%s-%02d", year, month)
      combined <- NULL
      for (depth in c("-55.0", "-105.0", "-280.0")) {
        added <- paste0("avg_depth_", depth, "_month_", date, ".csv")
        temp_path <- file.path(choose_folder("temperature", year), paste0("temperature_", added))
        sal_path  <- file.path(choose_folder("salinity", year), paste0("salinity_", added))
        pres_path <- file.path(choose_folder("pressure", year), paste0("pressure_", added))
        mvel_path <- file.path(choose_folder("mvel", year), paste0("meridional_velocity_", added))
        zvel_path <- file.path(choose_folder("zvel", year), paste0("zonal_velocity_", added))

        if (!file.exists(temp_path) || !file.exists(sal_path) || !file.exists(pres_path)) next

        depth_clean <- str_remove_all(depth, "[-]|\\.0$")
        temp_var <- paste0("temperature_depth_", depth_clean, "_std")
        sal_var  <- paste0("salinity_depth_", depth_clean, "_std")
        pres_var <- paste0("pressure_depth_", depth_clean, "_std")
        mvel_var <- paste0("meridional_velocity_depth_", depth_clean, "_std")
        zvel_var <- paste0("zonal_velocity_depth_", depth_clean, "_std")

        temp_long <- pivot_grid(read_csv(temp_path, show_col_types = FALSE), temp_var)
        sal_long  <- pivot_grid(read_csv(sal_path, show_col_types = FALSE),  sal_var)
        pres_long <- pivot_grid(read_csv(pres_path, show_col_types = FALSE), pres_var)
        mvel_long <- pivot_grid(read_csv(mvel_path, show_col_types = FALSE), mvel_var)
        zvel_long <- pivot_grid(read_csv(zvel_path, show_col_types = FALSE), zvel_var)

        depth_df <- temp_long %>%
          left_join(sal_long, by = c("lat", "lon")) %>%
          left_join(pres_long, by = c("lat", "lon")) %>%
          left_join(mvel_long, by = c("lat", "lon")) %>%
          left_join(zvel_long, by = c("lat", "lon"))

        combined <- if (is.null(combined)) depth_df else left_join(combined, depth_df, by = c("lat", "lon"))
      }
      for (depth in c("55", "105", "280")) {
        z_col <- paste0("zonal_velocity_depth_", depth, "_std")
        m_col <- paste0("meridional_velocity_depth_", depth, "_std")
        mag_col <- paste0("magnitude_depth_", depth, "_std")
        theta_col <- paste0("theta_depth_", depth, "_std")
        combined[[mag_col]] <- sqrt(combined[[z_col]]^2 + combined[[m_col]]^2)
        combined[[theta_col]] <- atan2(combined[[z_col]], combined[[m_col]])
      }
      for (depth in c("55", "105", "280")) {
        for (suffix in c("magnitude", "theta")) {
          col <- paste0(suffix, "_depth_", depth, "_std")
          combined[[col]] <- scale(combined[[col]])[, 1]
        }
      }
      combined <- combined %>% add_utm_columns(ll_names = c("lon", "lat")) %>% drop_na()
      season_data[[length(season_data) + 1]] <- combined
    }

    
    
    if (length(season_data) == 0) next

    combined_season <- bind_rows(season_data) %>%
      group_by(lon, lat) %>%
      summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop") %>%
      mutate(season = as.numeric(year) + season_val)
    pred <- predict(fit, newdata = combined_season, type = "response")
    pred$lon <- combined_season$lon
    pred$lat <- combined_season$lat
    observed <- test_data %>%
      filter(season == as.numeric(year) + season_val, !is.na(resp), resp > 0) %>%
      select(lon, lat, resp)
    
    # Define bounding box from prediction
    lon_min <- min(pred$lon, na.rm = TRUE)
    lon_max <- max(pred$lon, na.rm = TRUE)
    lat_min <- min(pred$lat, na.rm = TRUE)
    lat_max <- max(pred$lat, na.rm = TRUE)


    # Crop coastline to match prediction extent
    coast_crop <- sf::st_crop(coast, xmin = lon_min, xmax = lon_max,
                                  ymin = lat_min, ymax = lat_max)
    season_label <- c("Winter", "Spring", "Summer", "Fall")[[which(c(0, 0.25, 0.5, 0.75) == season_val)]]
    
    p <- ggplot(pred, aes(x = lon, y = lat)) +
    geom_tile(aes(fill = est)) +
    geom_sf(data = coast_crop, inherit.aes = FALSE, color = "black", fill = NA, size = 0.3) +
    coord_sf(xlim = c(lon_min, lon_max), ylim = c(lat_min, lat_max), expand = FALSE) +
    scale_x_continuous(name = "Longitude", labels = scales::number_format(accuracy = 1)) +
    scale_y_continuous(name = "Latitude", labels = scales::number_format(accuracy = 1)) +
    scale_fill_viridis_c(
      name = "Predicted",
      limits = if (cap > 0) c(0, cap) else NULL,
      oob = scales::squish
    ) +
    labs(
      title = paste(season_label, year)
    ) +
    theme_minimal() +
    theme(
      legend.background = element_rect(fill = "transparent", color = NA),
      legend.box.background = element_rect(fill = "transparent", color = NA),
      legend.key = element_rect(fill = "transparent", color = NA),
      panel.background = element_rect(fill = "transparent", color = NA)
    )

      
    # Add observed detections and non-detections
    if (nrow(observed) > 0) {
      observed <- observed %>%
        filter(lon >= lon_min, lon <= lon_max, lat >= lat_min, lat <= lat_max)

      observed[[resp]] <- as.numeric(as.character(observed[[resp]]))
      detected <- observed %>% filter(.data[[resp]] > 0)
      nondetected <- observed %>% filter(.data[[resp]] == 0)

      # Add detections (sized red points)
      if (nrow(detected) > 0) {
        p <- p +
          geom_point(
            data = detected,
            aes(x = lon, y = lat, size = .data[[resp]]),
            color = "red", shape = 21, stroke = 0.5, fill = "red"
          )
      }
    
      # Add non-detections (small blue dots)
      if (nrow(nondetected) > 0) {
        p <- p +
          geom_point(
            data = nondetected,
            aes(x = lon, y = lat),
            color = "lightblue", size = 0.5, shape = 21, stroke = 0.3, fill = "lightblue"
          )
      }
    max_obs <- max(test_data[[resp]], na.rm = TRUE)
    p <- p + scale_size_continuous(
      name = "Observed",
      range = c(1, 3),  # dot sizes (not legend) still start from 1
      limits = c(1, max_obs),  # includes 0 for completeness
      breaks = pretty(c(1, max_obs), n = 4)  # legend starts at 1
    )
  }


    plots[[season_label]] <- p
  }
  return(plots)
}
```

## Modeling

#### Intercept with iid

```{r}
# Create mesh using training data
mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 5)

# Fit spatiotemporal model
fit_spatiotemporal <- sdmTMB(
  d_scaled ~ 1, 
  data = train_data, 
  mesh = mesh_train,
  time = "season",
  family = delta_gamma(link1 = "logit", link2 = "log"), 
  spatial = "on", 
  spatiotemporal = "iid",
  extra_time = unique(test_data$season)
)

# Extract coefficients and sanity check
# tidy(fit_spatiotemporal, conf.int = TRUE)
sanity(fit_spatiotemporal)

# Predictions on training and test data
training_pred <- predict(fit_spatiotemporal, newdata = train_data, type = "response")
training_pred$observed <- train_data$d_scaled

test_pred <- predict(fit_spatiotemporal, newdata = test_data, type = "response")
test_pred$observed <- test_data$d_scaled

# Plotting the predictions
p0 <- ggplot(training_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Train") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

p1 <- ggplot(test_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Test") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

gridExtra::grid.arrange(p0, p1)

# Compute RMSE
rmse_train <- rmse(training_pred$observed, training_pred$est)
rmse_test <- rmse(test_pred$observed, test_pred$est)

# Compute Range
range_train <- range(train_data$d_scaled)
range_test <- range(test_data$d_scaled)

# Compute Standard Deviation
sd_train <- sd(train_data$d_scaled, na.rm = TRUE)
sd_test <- sd(test_data$d_scaled, na.rm = TRUE)

# Assemble the table
summary_table <- data.frame(
  Dataset = c("Training", "Test"),
  RMSE = c(rmse_train, rmse_test),
  Range_Min = c(range_train[1], range_test[1]),
  Range_Max = c(range_train[2], range_test[2]),
  Std_Dev = c(sd_train, sd_test)
)

# Print the summary table
print(summary_table)
```

```{r, include=FALSE}
d_plots <- generate_seasonal_prediction_plots(fit_spatiotemporal, year = "2023", test_data = test_data, cap = 0.5, resp = "d_scaled")
wrap_d_plots <- wrap_plots(d_plots, ncol = 2, nrow = 2, guides = "collect") +
  plot_annotation(
    title = "Seasonal D Call Density Predictions (2023)",
    theme = theme(plot.title = element_text(hjust = 0.5, size = 16))
  )
print(wrap_d_plots)
```


#### Intercept with rw

```{r}
# Create mesh using training data
mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 5)

# Fit spatiotemporal model
fit_spatiotemporal <- sdmTMB(
  d_scaled ~ 1, 
  data = train_data, 
  mesh = mesh_train,
  time = "season",
  family = delta_gamma(link1 = "logit", link2 = "log"), 
  spatial = "on", 
  spatiotemporal = "rw",
  extra_time = unique(test_data$season)
)

# Extract coefficients and sanity check
# tidy(fit_spatiotemporal, conf.int = TRUE)
sanity(fit_spatiotemporal)

# Predictions on training and test data
training_pred <- predict(fit_spatiotemporal, newdata = train_data, type = "response")
training_pred$observed <- train_data$d_scaled

test_pred <- predict(fit_spatiotemporal, newdata = test_data, type = "response")
test_pred$observed <- test_data$d_scaled

# Plotting the predictions
p0 <- ggplot(training_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Train") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

p1 <- ggplot(test_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Test") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

gridExtra::grid.arrange(p0, p1)

# Compute RMSE
rmse_train <- rmse(training_pred$observed, training_pred$est)
rmse_test <- rmse(test_pred$observed, test_pred$est)

# Compute Range
range_train <- range(train_data$d_scaled)
range_test <- range(test_data$d_scaled)

# Compute Standard Deviation
sd_train <- sd(train_data$d_scaled, na.rm = TRUE)
sd_test <- sd(test_data$d_scaled, na.rm = TRUE)

# Assemble the table
summary_table <- data.frame(
  Dataset = c("Training", "Test"),
  RMSE = c(rmse_train, rmse_test),
  Range_Min = c(range_train[1], range_test[1]),
  Range_Max = c(range_train[2], range_test[2]),
  Std_Dev = c(sd_train, sd_test)
)

# Print the summary table
print(summary_table)

#d_plots <- generate_seasonal_prediction_plots(fit_spatiotemporal, year = "2023", test_data = #test_data, cap = 0.5, resp = "d_scaled")
#wrap_d_plots <- wrap_plots(d_plots, ncol = 2, nrow = 2, guides = "collect") +
#  plot_annotation(
#    title = "Seasonal D Call Density Predictions (2023)",
#    theme = theme(plot.title = element_text(hjust = 0.5, size = 16))
#  )
#print(wrap_d_plots)
```

#### Intercept with ar1

```{r}
# Create mesh using training data
mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 5)

# Fit spatiotemporal model
fit_spatiotemporal <- sdmTMB(
  d_scaled ~ 1, 
  data = train_data, 
  mesh = mesh_train,
  time = "season",
  family = delta_gamma(link1 = "logit", link2 = "log"), 
  spatial = "on", 
  spatiotemporal = "ar1",
  extra_time = unique(test_data$season)
)

# Extract coefficients and sanity check
# tidy(fit_spatiotemporal, conf.int = TRUE)
sanity(fit_spatiotemporal)

# Predictions on training and test data
training_pred <- predict(fit_spatiotemporal, newdata = train_data, type = "response")
training_pred$observed <- train_data$d_scaled

test_pred <- predict(fit_spatiotemporal, newdata = test_data, type = "response")
test_pred$observed <- test_data$d_scaled

# Plotting the predictions
p0 <- ggplot(training_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Train") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

p1 <- ggplot(test_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Test") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

gridExtra::grid.arrange(p0, p1)

# Compute RMSE
rmse_train <- rmse(training_pred$observed, training_pred$est)
rmse_test <- rmse(test_pred$observed, test_pred$est)

# Compute Range
range_train <- range(train_data$d_scaled)
range_test <- range(test_data$d_scaled)

# Compute Standard Deviation
sd_train <- sd(train_data$d_scaled, na.rm = TRUE)
sd_test <- sd(test_data$d_scaled, na.rm = TRUE)

# Assemble the table
summary_table <- data.frame(
  Dataset = c("Training", "Test"),
  RMSE = c(rmse_train, rmse_test),
  Range_Min = c(range_train[1], range_test[1]),
  Range_Max = c(range_train[2], range_test[2]),
  Std_Dev = c(sd_train, sd_test)
)

# Print the summary table
print(summary_table)

#d_plots <- generate_seasonal_prediction_plots(fit_spatiotemporal, year = "2023", test_data = #test_data, cap = 0.5, resp = "d_scaled")
#wrap_d_plots <- wrap_plots(d_plots, ncol = 2, nrow = 2, guides = "collect") +
#  plot_annotation(
#    title = "Seasonal D Call Density Predictions (2023)",
#    theme = theme(plot.title = element_text(hjust = 0.5, size = 16))
#  )
#print(wrap_d_plots)
```


#### All Environmental Predictors with iid

```{r}
# Create mesh using training data
mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 5)

# Fit spatiotemporal model
fit_spatiotemporal <- sdmTMB(
  d_scaled ~ temperature_depth_105_std + temperature_depth_55_std + temperature_depth_280_std + salinity_depth_55_std + salinity_depth_105_std + magnitude_depth_55_std + magnitude_depth_105_std + theta_depth_55_std + theta_depth_105_std,  
  data = train_data, 
  mesh = mesh_train,
  time = "season",
  family = delta_gamma(link1 = "logit", link2 = "log"), 
  spatial = "on", 
  spatiotemporal = "iid",
  extra_time = unique(test_data$season)
)

# Extract coefficients and sanity check
#tidy(fit_spatiotemporal, conf.int = TRUE)
sanity(fit_spatiotemporal)

# Predictions on training and test data
training_pred <- predict(fit_spatiotemporal, newdata = train_data, type = "response")
training_pred$observed <- train_data$d_scaled

test_pred <- predict(fit_spatiotemporal, newdata = test_data, type = "response")
test_pred$observed <- test_data$d_scaled

# Plotting the predictions
p0 <- ggplot(training_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Train") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

p1 <- ggplot(test_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Test") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

gridExtra::grid.arrange(p0, p1)

# Compute RMSE
rmse_train <- rmse(training_pred$observed, training_pred$est)
rmse_test <- rmse(test_pred$observed, test_pred$est)

# Compute Range
range_train <- range(train_data$d_scaled, na.rm = TRUE)
range_test <- range(test_data$d_scaled, na.rm = TRUE)

# Compute Standard Deviation
sd_train <- sd(train_data$d_scaled, na.rm = TRUE)
sd_test <- sd(test_data$d_scaled, na.rm = TRUE)

# Assemble the table
summary_table <- data.frame(
  Dataset = c("Training", "Test"),
  RMSE = c(rmse_train, rmse_test),
  Range_Min = c(range_train[1], range_test[1]),
  Range_Max = c(range_train[2], range_test[2]),
  Std_Dev = c(sd_train, sd_test)
)

# Print the summary table
print(summary_table)
```

#### All Environmental Predictors with rw

```{r}
# Create mesh using training data
mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 5)

# Fit spatiotemporal model
fit_spatiotemporal <- sdmTMB(
  d_scaled ~ temperature_depth_105_std + temperature_depth_55_std + temperature_depth_280_std + salinity_depth_55_std + salinity_depth_105_std + magnitude_depth_55_std + magnitude_depth_105_std + theta_depth_55_std + theta_depth_105_std,  
  data = train_data, 
  mesh = mesh_train,
  time = "season",
  family = delta_gamma(link1 = "logit", link2 = "log"), 
  spatial = "on", 
  spatiotemporal = "rw",
  extra_time = unique(test_data$season)
)

# Extract coefficients and sanity check
#tidy(fit_spatiotemporal, conf.int = TRUE)
sanity(fit_spatiotemporal)

# Predictions on training and test data
training_pred <- predict(fit_spatiotemporal, newdata = train_data, type = "response")
training_pred$observed <- train_data$d_scaled

test_pred <- predict(fit_spatiotemporal, newdata = test_data, type = "response")
test_pred$observed <- test_data$d_scaled

# Plotting the predictions
p0 <- ggplot(training_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Train") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

p1 <- ggplot(test_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Test") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

gridExtra::grid.arrange(p0, p1)

# Compute RMSE
rmse_train <- rmse(training_pred$observed, training_pred$est)
rmse_test <- rmse(test_pred$observed, test_pred$est)

# Compute Range
range_train <- range(train_data$d_scaled, na.rm = TRUE)
range_test <- range(test_data$d_scaled, na.rm = TRUE)

# Compute Standard Deviation
sd_train <- sd(train_data$d_scaled, na.rm = TRUE)
sd_test <- sd(test_data$d_scaled, na.rm = TRUE)

# Assemble the table
summary_table <- data.frame(
  Dataset = c("Training", "Test"),
  RMSE = c(rmse_train, rmse_test),
  Range_Min = c(range_train[1], range_test[1]),
  Range_Max = c(range_train[2], range_test[2]),
  Std_Dev = c(sd_train, sd_test)
)

# Print the summary table
print(summary_table)

```

#### All Environmental Predictors with ar1

```{r}
# Create mesh using training data
mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 5)

# Fit spatiotemporal model
fit_spatiotemporal <- sdmTMB(
  d_scaled ~ temperature_depth_105_std + temperature_depth_55_std + temperature_depth_280_std + salinity_depth_55_std + salinity_depth_105_std + magnitude_depth_55_std + magnitude_depth_105_std + theta_depth_55_std + theta_depth_105_std,  
  data = train_data, 
  mesh = mesh_train,
  time = "season",
  family = delta_gamma(link1 = "logit", link2 = "log"), 
  spatial = "on", 
  spatiotemporal = "ar1",
  extra_time = unique(test_data$season)
)

# Extract coefficients and sanity check
#tidy(fit_spatiotemporal, conf.int = TRUE)
sanity(fit_spatiotemporal)

# Predictions on training and test data
training_pred <- predict(fit_spatiotemporal, newdata = train_data, type = "response")
training_pred$observed <- train_data$d_scaled

test_pred <- predict(fit_spatiotemporal, newdata = test_data, type = "response")
test_pred$observed <- test_data$d_scaled

# Plotting the predictions
p0 <- ggplot(training_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Train") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

p1 <- ggplot(test_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Test") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

gridExtra::grid.arrange(p0, p1)

# Compute RMSE
rmse_train <- rmse(training_pred$observed, training_pred$est)
rmse_test <- rmse(test_pred$observed, test_pred$est)

# Compute Range
range_train <- range(train_data$d_scaled, na.rm = TRUE)
range_test <- range(test_data$d_scaled, na.rm = TRUE)

# Compute Standard Deviation
sd_train <- sd(train_data$d_scaled, na.rm = TRUE)
sd_test <- sd(test_data$d_scaled, na.rm = TRUE)

# Assemble the table
summary_table <- data.frame(
  Dataset = c("Training", "Test"),
  RMSE = c(rmse_train, rmse_test),
  Range_Min = c(range_train[1], range_test[1]),
  Range_Max = c(range_train[2], range_test[2]),
  Std_Dev = c(sd_train, sd_test)
)

# Print the summary table
print(summary_table)

```

#### depth55 with iid

```{r}
# Create mesh using training data
mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 5)

# Fit spatiotemporal model
fit_spatiotemporal <- sdmTMB(
  d_scaled ~ temperature_depth_55_std + salinity_depth_55_std + magnitude_depth_55_std + theta_depth_55_std,  
  data = train_data, 
  mesh = mesh_train,
  time = "season",
  family = delta_gamma(link1 = "cloglog", link2 = "log"), 
  spatial = "on", 
  spatiotemporal = "iid",
  extra_time = unique(test_data$season)
)

# Extract coefficients and sanity check
#tidy(fit_spatiotemporal, conf.int = TRUE)
sanity(fit_spatiotemporal)

# Predictions on training and test data
training_pred <- predict(fit_spatiotemporal, newdata = train_data, type = "response")
training_pred$observed <- train_data$d_scaled

test_pred <- predict(fit_spatiotemporal, newdata = test_data, type = "response")
test_pred$observed <- test_data$d_scaled

# Plotting the predictions
p0 <- ggplot(training_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Train") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

p1 <- ggplot(test_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Test") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

gridExtra::grid.arrange(p0, p1)

# Compute RMSE
rmse_train <- rmse(training_pred$observed, training_pred$est)
rmse_test <- rmse(test_pred$observed, test_pred$est)

# Compute Range
range_train <- range(train_data$d_scaled, na.rm = TRUE)
range_test <- range(test_data$d_scaled, na.rm = TRUE)

# Compute Standard Deviation
sd_train <- sd(train_data$d_scaled, na.rm = TRUE)
sd_test <- sd(test_data$d_scaled, na.rm = TRUE)

# Assemble the table
summary_table <- data.frame(
  Dataset = c("Training", "Test"),
  RMSE = c(rmse_train, rmse_test),
  Range_Min = c(range_train[1], range_test[1]),
  Range_Max = c(range_train[2], range_test[2]),
  Std_Dev = c(sd_train, sd_test)
)

# Print the summary table
print(summary_table)
```

#### depth55 with rw

```{r}
# Create mesh using training data
mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 5)

# Fit spatiotemporal model
fit_spatiotemporal <- sdmTMB(
  d_scaled ~ temperature_depth_55_std + salinity_depth_55_std + magnitude_depth_55_std + theta_depth_55_std,  
  data = train_data, 
  mesh = mesh_train,
  time = "season",
  family = delta_gamma(link1 = "cloglog", link2 = "log"), 
  spatial = "on", 
  spatiotemporal = "rw",
  extra_time = unique(test_data$season)
)

# Extract coefficients and sanity check
#tidy(fit_spatiotemporal, conf.int = TRUE)
sanity(fit_spatiotemporal)

# Predictions on training and test data
training_pred <- predict(fit_spatiotemporal, newdata = train_data, type = "response")
training_pred$observed <- train_data$d_scaled

test_pred <- predict(fit_spatiotemporal, newdata = test_data, type = "response")
test_pred$observed <- test_data$d_scaled

# Plotting the predictions
p0 <- ggplot(training_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Train") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

p1 <- ggplot(test_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Test") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

gridExtra::grid.arrange(p0, p1)

# Compute RMSE
rmse_train <- rmse(training_pred$observed, training_pred$est)
rmse_test <- rmse(test_pred$observed, test_pred$est)

# Compute Range
range_train <- range(train_data$d_scaled, na.rm = TRUE)
range_test <- range(test_data$d_scaled, na.rm = TRUE)

# Compute Standard Deviation
sd_train <- sd(train_data$d_scaled, na.rm = TRUE)
sd_test <- sd(test_data$d_scaled, na.rm = TRUE)

# Assemble the table
summary_table <- data.frame(
  Dataset = c("Training", "Test"),
  RMSE = c(rmse_train, rmse_test),
  Range_Min = c(range_train[1], range_test[1]),
  Range_Max = c(range_train[2], range_test[2]),
  Std_Dev = c(sd_train, sd_test)
)

# Print the summary table
print(summary_table)

```

#### depth55 with ar1

```{r}
# Create mesh using training data
mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 5)

# Fit spatiotemporal model
fit_spatiotemporal <- sdmTMB(
  d_scaled ~ temperature_depth_55_std + salinity_depth_55_std + magnitude_depth_55_std + theta_depth_55_std,  
  data = train_data, 
  mesh = mesh_train,
  time = "season",
  family = delta_gamma(link1 = "cloglog", link2 = "log"), 
  spatial = "on", 
  spatiotemporal = "ar1",
  extra_time = unique(test_data$season)
)

# Extract coefficients and sanity check
#tidy(fit_spatiotemporal, conf.int = TRUE)
sanity(fit_spatiotemporal)

# Predictions on training and test data
training_pred <- predict(fit_spatiotemporal, newdata = train_data, type = "response")
training_pred$observed <- train_data$d_scaled

test_pred <- predict(fit_spatiotemporal, newdata = test_data, type = "response")
test_pred$observed <- test_data$d_scaled

# Plotting the predictions
p0 <- ggplot(training_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Train") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

p1 <- ggplot(test_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Test") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

gridExtra::grid.arrange(p0, p1)

# Compute RMSE
rmse_train <- rmse(training_pred$observed, training_pred$est)
rmse_test <- rmse(test_pred$observed, test_pred$est)

# Compute Range
range_train <- range(train_data$d_scaled, na.rm = TRUE)
range_test <- range(test_data$d_scaled, na.rm = TRUE)

# Compute Standard Deviation
sd_train <- sd(train_data$d_scaled, na.rm = TRUE)
sd_test <- sd(test_data$d_scaled, na.rm = TRUE)

# Assemble the table
summary_table <- data.frame(
  Dataset = c("Training", "Test"),
  RMSE = c(rmse_train, rmse_test),
  Range_Min = c(range_train[1], range_test[1]),
  Range_Max = c(range_train[2], range_test[2]),
  Std_Dev = c(sd_train, sd_test)
)

# Print the summary table
print(summary_table)

```

#### tem&salinity with iid

```{r}
# Create mesh using training data
mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 5)

# Fit spatiotemporal model
fit_spatiotemporal <- sdmTMB(
  d_scaled ~ temperature_depth_55_std + salinity_depth_55_std,
  data = train_data, 
  mesh = mesh_train,
  time = "season",
  family = delta_gamma(link1 = "cloglog", link2 = "log"), 
  spatial = "on", 
  spatiotemporal = "iid",
  extra_time = unique(test_data$season)
)

# Extract coefficients and sanity check
#tidy(fit_spatiotemporal, conf.int = TRUE)
sanity(fit_spatiotemporal)

# Predictions on training and test data
training_pred <- predict(fit_spatiotemporal, newdata = train_data, type = "response")
training_pred$observed <- train_data$d_scaled

test_pred <- predict(fit_spatiotemporal, newdata = test_data, type = "response")
test_pred$observed <- test_data$d_scaled

# Plotting the predictions
p0 <- ggplot(training_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Train") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

p1 <- ggplot(test_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Test") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

gridExtra::grid.arrange(p0, p1)

# Compute RMSE
rmse_train <- rmse(training_pred$observed, training_pred$est)
rmse_test <- rmse(test_pred$observed, test_pred$est)

# Compute Range
range_train <- range(train_data$d_scaled, na.rm = TRUE)
range_test <- range(test_data$d_scaled, na.rm = TRUE)

# Compute Standard Deviation
sd_train <- sd(train_data$d_scaled, na.rm = TRUE)
sd_test <- sd(test_data$d_scaled, na.rm = TRUE)

# Assemble the table
summary_table <- data.frame(
  Dataset = c("Training", "Test"),
  RMSE = c(rmse_train, rmse_test),
  Range_Min = c(range_train[1], range_test[1]),
  Range_Max = c(range_train[2], range_test[2]),
  Std_Dev = c(sd_train, sd_test)
)

# Print the summary table
print(summary_table)


```

#### tem&salinity with rw

```{r}
# Create mesh using training data
mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 5)

# Fit spatiotemporal model
fit_spatiotemporal <- sdmTMB(
  d_scaled ~ temperature_depth_55_std + salinity_depth_55_std,
  data = train_data, 
  mesh = mesh_train,
  time = "season",
  family = delta_gamma(link1 = "cloglog", link2 = "log"), 
  spatial = "on", 
  spatiotemporal = "rw",
  extra_time = unique(test_data$season)
)

# Extract coefficients and sanity check
#tidy(fit_spatiotemporal, conf.int = TRUE)
sanity(fit_spatiotemporal)

# Predictions on training and test data
training_pred <- predict(fit_spatiotemporal, newdata = train_data, type = "response")
training_pred$observed <- train_data$d_scaled

test_pred <- predict(fit_spatiotemporal, newdata = test_data, type = "response")
test_pred$observed <- test_data$d_scaled

# Plotting the predictions
p0 <- ggplot(training_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Train") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

p1 <- ggplot(test_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Test") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

gridExtra::grid.arrange(p0, p1)

# Compute RMSE
rmse_train <- rmse(training_pred$observed, training_pred$est)
rmse_test <- rmse(test_pred$observed, test_pred$est)

# Compute Range
range_train <- range(train_data$d_scaled, na.rm = TRUE)
range_test <- range(test_data$d_scaled, na.rm = TRUE)

# Compute Standard Deviation
sd_train <- sd(train_data$d_scaled, na.rm = TRUE)
sd_test <- sd(test_data$d_scaled, na.rm = TRUE)

# Assemble the table
summary_table <- data.frame(
  Dataset = c("Training", "Test"),
  RMSE = c(rmse_train, rmse_test),
  Range_Min = c(range_train[1], range_test[1]),
  Range_Max = c(range_train[2], range_test[2]),
  Std_Dev = c(sd_train, sd_test)
)

# Print the summary table
print(summary_table)

```

#### tem&salinity with ar1

```{r}
# Create mesh using training data
mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 5)

# Fit spatiotemporal model
fit_spatiotemporal <- sdmTMB(
  d_scaled ~ temperature_depth_55_std + salinity_depth_55_std,
  data = train_data, 
  mesh = mesh_train,
  time = "season",
  family = delta_gamma(link1 = "cloglog", link2 = "log"), 
  spatial = "on", 
  spatiotemporal = "ar1",
  extra_time = unique(test_data$season)
)

# Extract coefficients and sanity check
#tidy(fit_spatiotemporal, conf.int = TRUE)
sanity(fit_spatiotemporal)

# Predictions on training and test data
training_pred <- predict(fit_spatiotemporal, newdata = train_data, type = "response")
training_pred$observed <- train_data$d_scaled

test_pred <- predict(fit_spatiotemporal, newdata = test_data, type = "response")
test_pred$observed <- test_data$d_scaled

# Plotting the predictions
p0 <- ggplot(training_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Train") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

p1 <- ggplot(test_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Test") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

gridExtra::grid.arrange(p0, p1)

# Compute RMSE
rmse_train <- rmse(training_pred$observed, training_pred$est)
rmse_test <- rmse(test_pred$observed, test_pred$est)

# Compute Range
range_train <- range(train_data$d_scaled, na.rm = TRUE)
range_test <- range(test_data$d_scaled, na.rm = TRUE)

# Compute Standard Deviation
sd_train <- sd(train_data$d_scaled, na.rm = TRUE)
sd_test <- sd(test_data$d_scaled, na.rm = TRUE)

# Assemble the table
summary_table <- data.frame(
  Dataset = c("Training", "Test"),
  RMSE = c(rmse_train, rmse_test),
  Range_Min = c(range_train[1], range_test[1]),
  Range_Max = c(range_train[2], range_test[2]),
  Std_Dev = c(sd_train, sd_test)
)

# Print the summary table
print(summary_table)

```

#### tem&mag with iid

```{r}
# Create mesh using training data
mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 5)

# Fit spatiotemporal model
fit_spatiotemporal <- sdmTMB(
  d_scaled ~ temperature_depth_55_std + magnitude_depth_55_std,
  data = train_data, 
  mesh = mesh_train,
  time = "season",
  family = delta_gamma(link1 = "logit", link2 = "log"), 
  spatial = "on", 
  spatiotemporal = "iid",
  extra_time = unique(test_data$season)
)

# Extract coefficients and sanity check
#tidy(fit_spatiotemporal, conf.int = TRUE)
sanity(fit_spatiotemporal)

# Predictions on training and test data
training_pred <- predict(fit_spatiotemporal, newdata = train_data, type = "response")
training_pred$observed <- train_data$d_scaled

test_pred <- predict(fit_spatiotemporal, newdata = test_data, type = "response")
test_pred$observed <- test_data$d_scaled

# Plotting the predictions
p0 <- ggplot(training_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Train") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

p1 <- ggplot(test_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Test") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

gridExtra::grid.arrange(p0, p1)

# Compute RMSE
rmse_train <- rmse(training_pred$observed, training_pred$est)
rmse_test <- rmse(test_pred$observed, test_pred$est)

# Compute Range
range_train <- range(train_data$d_scaled, na.rm = TRUE)
range_test <- range(test_data$d_scaled, na.rm = TRUE)

# Compute Standard Deviation
sd_train <- sd(train_data$d_scaled, na.rm = TRUE)
sd_test <- sd(test_data$d_scaled, na.rm = TRUE)

# Assemble the table
summary_table <- data.frame(
  Dataset = c("Training", "Test"),
  RMSE = c(rmse_train, rmse_test),
  Range_Min = c(range_train[1], range_test[1]),
  Range_Max = c(range_train[2], range_test[2]),
  Std_Dev = c(sd_train, sd_test)
)

# Print the summary table
print(summary_table)

```

#### tem&mag with rw

```{r}
# Create mesh using training data
mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 5)

# Fit spatiotemporal model
fit_spatiotemporal <- sdmTMB(
  d_scaled ~ temperature_depth_55_std + magnitude_depth_55_std,
  data = train_data, 
  mesh = mesh_train,
  time = "season",
  family = delta_gamma(link1 = "logit", link2 = "log"), 
  spatial = "on", 
  spatiotemporal = "rw",
  extra_time = unique(test_data$season)
)

# Extract coefficients and sanity check
#tidy(fit_spatiotemporal, conf.int = TRUE)
sanity(fit_spatiotemporal)

# Predictions on training and test data
training_pred <- predict(fit_spatiotemporal, newdata = train_data, type = "response")
training_pred$observed <- train_data$d_scaled

test_pred <- predict(fit_spatiotemporal, newdata = test_data, type = "response")
test_pred$observed <- test_data$d_scaled

# Plotting the predictions
p0 <- ggplot(training_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Train") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

p1 <- ggplot(test_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Test") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

gridExtra::grid.arrange(p0, p1)

# Compute RMSE
rmse_train <- rmse(training_pred$observed, training_pred$est)
rmse_test <- rmse(test_pred$observed, test_pred$est)

# Compute Range
range_train <- range(train_data$d_scaled, na.rm = TRUE)
range_test <- range(test_data$d_scaled, na.rm = TRUE)

# Compute Standard Deviation
sd_train <- sd(train_data$d_scaled, na.rm = TRUE)
sd_test <- sd(test_data$d_scaled, na.rm = TRUE)

# Assemble the table
summary_table <- data.frame(
  Dataset = c("Training", "Test"),
  RMSE = c(rmse_train, rmse_test),
  Range_Min = c(range_train[1], range_test[1]),
  Range_Max = c(range_train[2], range_test[2]),
  Std_Dev = c(sd_train, sd_test)
)

# Print the summary table
print(summary_table)

```

#### tem&mag with ar1

```{r}
# Create mesh using training data
mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 5)

# Fit spatiotemporal model
fit_spatiotemporal <- sdmTMB(
  d_scaled ~ temperature_depth_55_std + magnitude_depth_55_std,
  data = train_data, 
  mesh = mesh_train,
  time = "season",
  family = delta_gamma(link1 = "logit", link2 = "log"), 
  spatial = "on", 
  spatiotemporal = "ar1",
  extra_time = unique(test_data$season)
)

# Extract coefficients and sanity check
#tidy(fit_spatiotemporal, conf.int = TRUE)
sanity(fit_spatiotemporal)

# Predictions on training and test data
training_pred <- predict(fit_spatiotemporal, newdata = train_data, type = "response")
training_pred$observed <- train_data$d_scaled

test_pred <- predict(fit_spatiotemporal, newdata = test_data, type = "response")
test_pred$observed <- test_data$d_scaled

# Plotting the predictions
p0 <- ggplot(training_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Train") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

p1 <- ggplot(test_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Test") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

gridExtra::grid.arrange(p0, p1)

# Compute RMSE
rmse_train <- rmse(training_pred$observed, training_pred$est)
rmse_test <- rmse(test_pred$observed, test_pred$est)

# Compute Range
range_train <- range(train_data$d_scaled, na.rm = TRUE)
range_test <- range(test_data$d_scaled, na.rm = TRUE)

# Compute Standard Deviation
sd_train <- sd(train_data$d_scaled, na.rm = TRUE)
sd_test <- sd(test_data$d_scaled, na.rm = TRUE)

# Assemble the table
summary_table <- data.frame(
  Dataset = c("Training", "Test"),
  RMSE = c(rmse_train, rmse_test),
  Range_Min = c(range_train[1], range_test[1]),
  Range_Max = c(range_train[2], range_test[2]),
  Std_Dev = c(sd_train, sd_test)
)

# Print the summary table
print(summary_table)

```

#### lon&lat&depth55 with iid

```{r}
# Create mesh using training data
mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 10)

# Fit spatiotemporal model
fit_spatiotemporal <- sdmTMB(
  d_scaled ~ lat + lon + temperature_depth_55_std + salinity_depth_55_std + magnitude_depth_55_std + theta_depth_55_std,  
  data = train_data, 
  mesh = mesh_train,
  time = "season",
  family = delta_gamma(link1 = "logit", link2 = "log"), 
  spatial = "on", 
  spatiotemporal = "iid",
  extra_time = unique(test_data$season)
)

# Extract coefficients and sanity check
#tidy(fit_spatiotemporal, conf.int = TRUE)
sanity(fit_spatiotemporal)

# Predictions on training and test data
training_pred <- predict(fit_spatiotemporal, newdata = train_data, type = "response")
training_pred$observed <- train_data$d_scaled

test_pred <- predict(fit_spatiotemporal, newdata = test_data, type = "response")
test_pred$observed <- test_data$d_scaled

# Plotting the predictions
p0 <- ggplot(training_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Train") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

p1 <- ggplot(test_pred, aes(x = observed, y = est)) + 
  geom_point() + 
  ggtitle("Test") + 
  geom_abline(intercept = 0, slope = 1, col = "red")

gridExtra::grid.arrange(p0, p1)

# Compute RMSE
rmse_train <- rmse(training_pred$observed, training_pred$est)
rmse_test <- rmse(test_pred$observed, test_pred$est)

# Compute Range
range_train <- range(train_data$d_scaled, na.rm = TRUE)
range_test <- range(test_data$d_scaled, na.rm = TRUE)

# Compute Standard Deviation
sd_train <- sd(train_data$d_scaled, na.rm = TRUE)
sd_test <- sd(test_data$d_scaled, na.rm = TRUE)

# Assemble the table
summary_table <- data.frame(
  Dataset = c("Training", "Test"),
  RMSE = c(rmse_train, rmse_test),
  Range_Min = c(range_train[1], range_test[1]),
  Range_Max = c(range_train[2], range_test[2]),
  Std_Dev = c(sd_train, sd_test)
)

# Print the summary table
print(summary_table)
```

```{r}
# Prediction heatmap
d_plots <- generate_seasonal_prediction_plots(fit_spatiotemporal, year = "2023", test_data = test_data, cap = 0.5, resp = "d_scaled")
wrap_d_plots <- wrap_plots(d_plots, ncol = 2, nrow = 2, guides = "collect") +
  plot_annotation(
    title = "Seasonal D Call Density Predictions (2023)",
    theme = theme(plot.title = element_text(hjust = 0.5, size = 16))
  )
print(wrap_d_plots)
```