---
title: "Stepwise Selection"
output: html_document
---

```{r, message = F, warning = F}
# Load in the libraries
library(tidyverse)
library(sdmTMB)
library(ggplot2)
library(janitor)
library(sp)
library(readr)
library(ggcorrplot)
library(Metrics)
library(naniar)
library(tibble)
library(rsample)
library(caret)
library(purrr)
library(tune)

# Read in the data
cal_data <- read.csv("/Users/josh/Documents/GitHub/capstone-scripps/data/Merged_CalCOFI/STD_CalCOFI_final")
```

```{r}
# Removing the first two columns
cal_data <- cal_data[, -c(1, 2)]

# Removing larvae and euphasia abundance
cal_data <- cal_data %>% 
  select(-euphasia_abundance, -euphasia_abundance_std, -larvae_count, -larvae_count_std, - total_abundance, -total_abundance_std)

# Changing dataset to only have present response data
cal_data <- cal_data %>% 
  filter(!is.na(b_scaled))
```


```{r}
cols_to_use <- c("est_depth_std","temperature_depth_105_std", "temperature_depth_55_std", "temperature_depth_280_std", "salinity_depth_55_std", "salinity_depth_105_std", "magnitude_depth_55_std", "magnitude_depth_105_std",  "theta_depth_55_std", "theta_depth_105_std", "lon", "lat", "season")

existing_cols <- intersect(cols_to_use, colnames(cal_data))
cor_data <- cal_data[, existing_cols, drop = FALSE]

# Compute correlation matrix (use complete.obs to skip missing data)
cor_matrix <- cor(cor_data, use = "complete.obs")

# Correlation plot
ggcorrplot(cor_matrix, 
           lab = TRUE, 
           hc.order = TRUE, 
           lab_size =2,
           tl.cex = 5,
           type = "lower", 
           title = "Correlation Plot")

```

```{r}

all_combos <- unlist(
  lapply(1:length(cols_to_use), function(i) {
    combn(cols_to_use, i, simplify = FALSE)
  }), recursive = FALSE
)
```


```{r}

set.seed(123)
last_year <- max(cal_data$year)
train_data <- cal_data %>% filter(year != last_year)
test_data <- cal_data %>% filter(year == last_year)

mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 5)

results <- map_dfr(all_combos, function(preds) {
  # build formula: if preds is empty use "~ 1", else list predictors
  rhs <- if (length(preds)) paste(preds, collapse = " + ") else "1"
  f <- as.formula(paste("b_scaled ~", rhs))
  
  # try fitting, capture errors
  fit <- tryCatch({
    sdmTMB(
      formula        = f,
      data           = train_data,
      mesh           = mesh_train,
      time           = "season",
      family         = delta_gamma(link1 = "logit", link2 = "log"),
      spatial        = "on",
      spatiotemporal = "rw",
      extra_time     = unique(test_data$season)
    )
  }, error = function(e) NULL)
  
  # extract AIC or mark as NA on failure
  tibble(
    predictors = paste(preds, collapse = ", "),
    AIC        = if (!is.null(fit)) AIC(fit) else NA_real_
  )
})

# 3. Find the combination with lowest AIC
best <- results %>% filter(!is.na(AIC)) %>% arrange(AIC)
best
```

```{r, eval = FALSE}
set.seed(123)
last_year <- max(cal_data$year)
train_data <- cal_data %>% filter(year != last_year)
test_data <- cal_data %>% filter(year == last_year)

mesh_train <- make_mesh(train_data,
                        xy_cols = c("X", "Y"),
                        cutoff = 10)

fit_spatiotemporal <- sdmTMB(
  b_scaled ~ est_depth_std,  
  data = train_data, 
  mesh = mesh_train,
  time = "season",
  family = delta_gamma(link1 = "logit", link2 = "log"), 
  spatial = "off", 
  spatiotemporal = "iid",
  extra_time = unique(test_data$season)
)

AIC(fit_spatiotemporal)
```

This is spatial on used temp 55, sal 55, lon, lat, est depth, mag 55, theta 55, lowkey forgot other 2, didnt take out NA first

A tibble: 244 × 2
    predictors                                                                AIC
    <chr>                                                                   <dbl>
1 est_depth_std                                                        -8.45e24
2 temperature_depth_55_std, salinity_depth_55_std, lon, lat             1.69e 3
3 est_depth_std, temperature_depth_55_std, salinity_depth_55_std, lat   1.69e 3
4 est_depth_std, temperature_depth_55_std, lat                          1.69e 3
5 est_depth_std, temperature_depth_55_std                               1.69e 3
6 est_depth_std, temperature_depth_55_std, salinity_depth_55_std, the…  1.69e 3
7 temperature_depth_55_std, lon, lat                                    1.69e 3
8 temperature_depth_55_std, salinity_depth_55_std, magnitude_depth_55…  1.69e 3
9 temperature_depth_55_std, salinity_depth_55_std, theta_depth_55_std…  1.69e 3
10 est_depth_std, temperature_depth_55_std, salinity_depth_55_std, mag…  1.69e 3
ℹ 234 more rows
